---
title: "Exploring gapminder"
author: "Your Name"
format: 
  html:
    fig-align: "center"
    fig-width: 4
    fig-height: 4
    fig-format: png
execute: 
  echo: true
  cache: false
---

```{r}
#| label: library
#| echo: true
#| message: false
library(tidyverse)
library(broom)
library(knitr)
```

# Introduction

Let's look at the gaminder data

```{r}
#| label: data-read-in
gapminder <- read_csv(here::here("data","gapminder.csv"))
gapminder_oz <- gapminder |> filter(country == "Australia")
```

let's look at the lifespan

```{r}
#| label: hist-life-exp
hist(gapminder$lifeExp)
```

Let's fit a simple linear model of the effect of year on life expectancy

```{r}
#| label: example-lm
fit <- lm(lifeExp ~ year, data = gapminder)
fit
```

And let's look at the coefficient table:

```{r}
#| label: coef-table
fit_coef <- tidy(fit)
knitr::kable(fit_coef,
             caption = "A table of the coefficients")
year_slope <- round(fit_coef$estimate[2], 3)
```

The effect of year on life expectancy is `r year_slope`.

# A table of the coefficients is:

```{r}
#| label: table-coef
knitr::kable(fit_coef)
```

```{r}
#| label: gg-oz-plot
#| fig-cap: Life expectancy from 1952 - 2007 for Australia. Life expentancy increases
#|   steadily except from 1962 to 1969. We can safely say that our life expectancy is
#|   higher than it has ever been!
library(ggplot2)
library(dplyr)
gapminder %>%
  filter(country == "Australia") %>%
  ggplot(aes(x = year,
             y = lifeExp)) + 
  geom_point()
```

Australia's life expectancy has increased a great deal over the past 50 years

(See Figure \@ref(fig:gg-oz-plot))

# References

